Experimental Analysis of a Spatialised Audio Interface for People with Visual Impairments

### 2020-05-21 paper.tex
Proofreading paper

\begin{abstract}
  Sound perception is a fundamental skill for many people with severe sight impairments.
  The research presented in this paper is part of an ongoing project with the aim to create a mobile guidance aid to help people with vision impairments find objects within an unknown indoor environment.
  This system requires an effective non-visual interface and uses bone-conduction headphones to transmit audio instructions to the user.
  It has been implemented and tested with spatialised audio cues, which convey the direction of a predefined target in 3D space.
  We present an in-depth evaluation of the audio interface with several experiments that involve a large number of participants, both blindfolded and with actual visual impairments, and analyse the pros and cons of our design choices.
  In addition to producing results comparable to the state-of-the-art, we found that Fitts's Law (a predictive model for human movement) can be used to generate a metric to improve and refine the quality of the audio interface in future mobile navigation aids. 
\end{abstract}

@@

\hl{The ActiVis project is an effort to create a fully mobile phone-based navigation aid for people with visual impairments to guide them towards a desired object or location within an unknown indoor environment.
Improvements to computer vision and machine learning techniques, as well as mobile computing hardware performance, are exploited to make this system possible.
In particular, techniques from the active vision field are used to enable a mobile device to gather information on the surrounding environment and use it to generate guidance instructions for a user with limited or no vision.
However, in the literature, these techniques are typically limited to directing electro-mechanical servos~\citep{bajcsy2018revisiting} and in this project, we attempt to find out if the same techniques can also be used to direct a user's attention to a target location. 
The proposed system was introduced by~\citet{lock2017portable}, who describe the aforementioned autonomous guidance system paired with a co-adaptive human-machine interface that changes its own parameters over time to better match the user's strengths and limitations.
In previous work we developed a prototype guidance system that uses active vision and machine learning models to gather information and help a person find objects within an unknown indoor environment, showing that these techniques can indeed successfully be applied to direct humans' attention~\citep{lock2019active}.
However, in this work we examine the effectiveness of the proposed interface for a searching task and investigate a metric that can be used in the next phases of the project to enable the aforementioned co-adaptive paradigm that could benefit the user experience and boost long-term navigation performance.}

The hardware used for the guidance system prototype are a Google Project Tango device\footnote{https://en.wikipedia.org/wiki/Tango\_(platform)}, along with a set of bone-conducting headphones (both are shown in \cref{fig:participant}).
This Tango device uses a set of embedded RGB-D cameras and other hardware and software components to provide accurate real-time localisation estimates and has powerful image-processing capabilities.
It also gives access to Android's full range of interface and I/O options.
Unlike normal headphones, bone-conducting headphones are placed on and conduct audio signals through a user's cheekbones, instead of their ear canal.
These headphones have the benefit of not interfering with a user's normal hearing function.

@@

The main contributions of this paper are two-fold: 
\begin{itemize}
  \item we provide comprehensive experimental results, with two groups of participants with healthy and limited eyesight, on how well a tone with varying pitch can convey a target's elevation angle when using a mobile device with a bone-conducting headset; 
  \item we show that this sound-based human-machine interface exhibits a response that is well-modelled by Fitts's Law which can provide a useful metric of performance for similar mobile user interfaces.
\end{itemize}

@@

\hl{Over the years, commercial and academic groups have devised new and innovative mobile navigation and electronic travel aids (ETA) for people with visual impairments to address macro- and micro-navigation issues.
  The latter refer to the scale that navigation takes place on, with macro-navigation addressing directional guidance on a topographical map using turn-by-turn instructions, for example, while micro-navigation focusses on conveying information of the user's immediate surroundings~\citep{petrie1997mobic}. 
Many of the proposed systems make use of one or a combination of vocal~\citep{mocanu2016when,chessa2016integrated,kanwal2015navigation,sato2019navcog}, audio~\citep{schwarze2015intuitive,rodriguez2012obstacle,kammoun2012navigation} and haptic~\citep{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to present a user with macro- and/or micro-guidance instructions, each of which has its own set of features and limitations.
The interface for this work is used to provide instructions to a stationary user to guide them to point a camera at a target location and we therefore are mainly concerned with micro-navigation tools. 
In this context, respondents with visual impairments typically prefer haptic and vocal feedback over an audio tone~\citep{arditi2013user,golledge2004stated}.
However, from an ergonomic perspective, haptic feedback has significant shortcomings and require additional external hardware to transmit guidance instructions with sufficient resolution.
Furthermore, in a micro-navigation task, high-resolution guidance and many adjustments are required to reach the target location.
This this regard, both haptic and vocal feedback can present a cognitive burden and overwhelm a user's input bandwidth, which could have a detrimental effect on performance and the user experience~\citep{klatzky2006cognitive}.}
%Indeed, \citet{arditi2013user} reports that users prefer controlling the rate at which guidance signals are presented to them, instead of being given constant guidance. 
As an alternative, simple audio tones are less prone to these bandwidth and hardware limitations.
However, such tones can potentially fatigue a user if there are too unpleasant.
Researchers have started using cameras and object detectors to determine what a user is looking at and then use simple audio tones and vocal feedback signals to indicate to the user where to find an object.
They report favourable results with these systems~\citep{schauerte2012assistive,tian2013computer,fiannaca2014headlock,vazquez2012helping}.

Work has been done in an attempt to spatialise these tones with a head-related transfer function (HRTF) that simulates a sound source being placed at some arbitrary 3D position.
The authors generally report favourable results when using normal over-ear headphones or speakers~\citep{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized,presti2019watchout,crispien1994guib}.
However, the audio transmission device used has a significant effect on performance.
Indeed, research has shown that cheaper headphones and bone-conducting headphones have diminished performance in location perception with HRTFs when compared to over-ear or other expensive headphones~\citep{stanley2006lateralization,voong2019auditory,mascetti2016sonification}.
However, this performance degradation seems limited to the elevation dimension, which is drastically improved when the HRTFs are adjusted for the bone-conduction pathway~\citep{stanley2006lateralization,pec2008individual}.
Another way around the performance issue in the elevation dimension is to transmit the elevation angle by adjusting the audio tone's pitch, such as in the work by~\citet{durette2008visuo}.
We expanded upon the latter's work and investigated using spatialised audio and a varying pitch for guidance in an initial study~\cite{lock2019bone}, and found that participants are able to adequately determine a target's elevation with this approach with performance in both pan and elevation dimensions.
The measured performance is comparable to those of more expensive and over-ear headphones.

Fitts's Law~\citep{fitts1954information} is a predictive model of human movement and is particularly useful to evaluate human-computer interactions.
Indeed, researchers have previously used Fitts's Law, and more recently MacKenzie's modified version of it~\citep{mackenzie1992fitts}, as a metric to evaluate the performance of a spatial audio HMI\@.
Fitts's Law was originally proposed for visual target search tasks, but has since been applied in non-visual target search tasks as well.
For example, experiments with a haptic feedback pointing device have been performed to evaluate how effective it was at directing a user towards a target~\citep{ahmaniemi2009augmented} and the authors showed that the search time adheres to Fitts's Law.
However, they also note that it is not a perfect fit, citing the fact that Fitts's Law does not take into account a user's search strategy.
Another group of researchers conducted experiments using a spatial audio interface to describe the position of a target on the horizontal plane~\citep{marentakis2006effects}.
Here, participants pointed to where they thought the targets were, on their left or right, as they traversed a path.
Their results show a good relation between target difficulty and search time, providing a strong argument that Fitts's Law can be used to describe the performance of a spatial audio interface.
These results have since been supported by other authors, who found that Fitts's Law provided a good explanation for the results from an experiment using visual, limited visual and non-visual feedback cues~\citep{wu2010fitts}.
However, Fitts's Law has not yet been shown to apply to a spatial tone that uses varying pitch to convey the target's elevation angle, as demonstrated in this paper.

@@

\hl{Existing electronic navigation aids have typically struggled to gain market traction and replace the traditional walking cane as the standard assistive tool for people with visual impairments.
Current technological limitations include prohibitive costs, bulky hardware requirements and non-user-friendly interfaces~\citep{golledge2004stated,yusif2016older,arditi2013user}.
To address these issues, we implemented a handheld mobile system that is based on a concept proposed by~\citet{lock2019active} and tested by~\citet{lock2019bone} using a Google Tango device that is able to localise itself in real-time.
This system has the benefit of minimal hardware requirements and a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.}

\hl{People with visual impairments rely heavily on their hearing~\citep{golledge2004stated} and we wish to avoid blocking this information pathway, since doing to can have severe undesirable effects~\citep{lichtenstein2012headphone}.
We therefore use a set of bone-conduction headphones that are placed on a user's cheekbones and conduct audio signals through the skull to the inner ear, completely avoiding the outer ear and therefore does not impede the user's ambient sound perception abilities. 
Alternative, open-back headphones were also considered, but it was found that they still interfere with hearing and were therefore disregarded.
The AfterShockz headphones (\cref{fig:participant}) were ultimately selected.
These headphones have the added benefit of a more discreet form-factor when compared to other over-headphones, thereby addressing the issue of user acceptance.}

@@

\hl{Humans are able to localise sounds in three dimensions by extracting different cues from the audio signal~\citep{blauert1969sound,blauert1997spatial}.
These include binaural, where the user compares the signals received at both ears (e.g.\ inter-aural time and level differences , ITD and ILD respectively) and monaural cues, where a cue is extracted directly from each ear's audio signal (e.g.\ the spectral profile, audio intensity).
The ITD and ILD extracted from binaural cues allow a user to localise a sound on the horizontal plane (e.g.\ a sound from the right will hit the right ear before left one), while monaural cues indicate a sound's distance (e.g. louder sounds appear closer than softer ones) and its position on the vertical plane (e.g.\ the spectral profile generated from the signal hitting the user's body).
Prior to transmitting an audio signal to the user, it can be adjusted by an HRTF to mimic a natural sound source and make a user believe it is located at some arbitrary 3D location. 
An HRTF is a mathematical function that simulates a human head and ear's response to an external sound and is derived by capturing key characteristics that affect the monaural and binaural responses at the user's ear, such as the hearing level and anatomy. 
An HRTF will produce the most accurate results when it is customised to match a specific user, since each user's hearing response is unique, but this is a complex process and HRTFs generated with average values (head sizes, heights, ear shapes, etc.) produce acceptable results~\citep{gardner1995hrtf}.}

\hl{The interface presents the user with guidance instructions in the form of angular adjustments that are required in the pan and elevation dimensions to point a camera at the desired target (see \cref{fig:participant}).
Spatialised audio is well-suited to this micro-navigation task, displaying accuracy levels comparable to vocal feedback, but with less cognitive load at the high resolution of this task~\citep{klatzky2006cognitive}. 
However, we propose a linear adjustment to the signal's pitch as a function of the elevation angle to overcome the limitations of bone-conduction headphones and spatialised audio stated previously.
The pan angle can still be conveyed with a spatialised audio signal generated by an HRTF.
Indeed, this dimension is unaffected by the use of bone-conduction headphones~\citep{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization,lock2019bone}. 
This interface was implemented and evaluated in~\citet{lock2019bone}.}

@@

\hl{To localise a sound on the horizontal plane, the human audition system compares characteristics from the signals received at both ears (binaural cues), such as their volume difference and the time delay between the same sound reaching both ears~\citep{blauert1969sound}.
For example, a sound placed at a person's right will hit their right ear first with a slightly higher volume compared to the left ear.
Since the binaural cues are largely independent if the signal's spectral profile, it would be convenient to use a simple audio wave to transmit guidance instructions to the user.
Therefore, in this work a pure sine wave was used, but can easily be replaced with a richer tone once the system's characteristics are better understood. 
The since was was spatialised using OpenAL's default HRTF , based on the MIT's KEMAR dataset~\citep{hiebert2005openal}.} %which uses the person and targets' positions as input, and outputs a transformed audio signal.}

@@

\hl{To compensate for the loss of elevation localisation performance when conveying spatialised audio via bone-conduction headphones~\citep{macdonald2006spatial,schonstein2008comparison}, we adjust the sine wave's frequency (i.e.\ audible pitch) as a linear function of the target's elevation angle, as shown in \cref{fig:pitch-gain}.
When the required elevation adjustment is above or below where the camera is currently pointed, the pitch is increased or lowered respectively. 
This high/low scheme was selected based on humans' natural association of high-pitched sounds with elevated objects and vice-versa for low-lying objects~\citep{pratt1930spatial,blauert1997spatial}.
The pitch is constantly adjusted and updated at \SI{10}{\hertz} in octave- and semitone-based intervals to ensure perceptible changes, while keeping the tone's timbre constant~\citep{shepard1964circularity}. }

\hl{The pitch is changed as a linear function of the elevation angle, the gradient of which is determined by setting the angle and pitch limits.
These limits are set at some number of octaves from the neutral pitch that is emitted when the camera is on-target. 
This tone is heuristically set to \SI{512}{\hertz} following practical tests, which is comfortably audible for a large number of octaves changes.
For this work, we only consider a \SI{180}{\degree} field of view in front of the user and limit the elevation angle to a range of \SI{\pm90}{\degree}, or $[-\frac{\pi}{2}, \frac{\pi}{2}]$.
In previous tests with this specific interface, its performance was comparable to normal over-ear headphones transmitting a spatialised signal that transmits the elevation angle~\citep{lock2019bone}.}

@@

We performed a set of experiments with the audio interface to determine how effective it is at directing a user to adjust the pan and elevation angles of a camera for point it to a target.
Furthermore, we also carried out a set of pre-screening experiments to determine each participant's hearing characteristics in order to determine their perception limits in the respective audio dimensions.
The participants were given time before the experiments commenced to familiarise themselves with the device and the tones it emits, as well as what the `on-target' tone sounds like.
We also tested the system with a group of participants with severe sight impairments and compared their data to the blindfolded participant dataset.
The results from the experiments we performed allow us to better understand how the users respond to different settings for the spatial audio feedback stimulus and use them to improve and optimise the behaviour of the feedback modes in our portable navigation aid.

@@

Even though the Tango is able to detect distances to object, we opted to use virtual targets given the added ability to place targets at any random location without physical manipulation while without losing the ability to determine its exact position relative to the device.
When this interface is used with a real navigation system, it will be extended to work with real targets.

@@

A set of experiments were conducted to determine the interface's guidance effectiveness for a pointing task.
These experiments captured the difference between the targets' actual directions and the directions the participants perceived them to be located. 
The participants were given a Tango tablet running an app implementing the experimental setup in \cref{fig:pipeline}.
This app generates a set of virtual targets, one at a time, and presents their directions to the participants with audio guidance signals.
These targets are generated at a constant distance from the participants with pan and elevation angles that are uniformly distributed across the four quadrants to avoid clustering.
Each target's relative angular position is communicated to the participants in real-time as the device is moved around via the bone-conducting headphones.
Every time a participant was confident that they were pointing at a target, i.e.\ hearing a signal that is placed at their front at a frequency of \SI{512}{\hertz}, they tapped the device's screen, marking the location and generating a new target.
The targets are all positioned relative to the device's camera coordinate system, which is tracked by the Tango hardware and localisation API.\
Each participant was tasked with finding 28 targets each.

@@

We also compare the performance of the three pitch gradient settings in terms of the time it takes each participant to find a target.
However, since each participant was presented with a different, randomly generated set of targets, a direct time comparison is not possible.
\hl{Instead, we use Fitts's Law~\citep{fitts1954information}, modified by MacKenzie for uncertain target sizes and noisy data~\citep{mackenzie1992fitts}, which states that there is a relation between the time it takes to find a target and its index of difficulty ($ID$, the ratio between the distance to the target and its width).
It also provides a so-called `index of performance' ($IP$), which can be used to compare the results between the three different configurations used in the experiments.
Furthermore, if a Fitts-like relationship is indeed found, it can be used as an optimisation metric in the co-adaptive interface described by~\cite{lock2017portable}, where interface parameters (e.g.\ pitch gradient, volume setting, etc.) are automatically tweaked in order to maximise target finding performance.}

@@

\noindent as proposed by~\citet{wobbrock2011effects}, taken as the angle between the participant's target selection and target's actual angular position.
\hl{The virtual targets have a programmed radius of approximately} \SI{10}{\centi\metre} (equivalent to approximately \SI{0.1}{\radian}) and when the participant pointed the camera within this radius, \SI{512}{\hertz} on-target tone was emitted.
\hl{This is similar to the work by}~\citet{kabbash1995prince}\hl{, where they had participants search for a point-target with an oversized cursor and found that this was similar to a Fitts-like response.
However, the participants were not explicitly informed when they were within this radius and had to use their own subjective judgement of the audio tone.
Since the judgement accuracy will vary across different participants, effectively giving different target sizes, it was necessary to use the effective width as an approximation for the actual target size.
Previous authors have found this to capture the true performance metrics fairly well}~\citep{zhai2004speed}.
Fitts's index of performance, $IP$, can then be calculated using

@@

Two groups of participants were recruited for the experiments on a volunteer basis. 
\hl{Group~\textit{G1} consisted of 42 undergraduate students (10 male, 32, female) with normal eyesight who were blindfolded for the experiments (mean age: $20\pm2$).
Group \textit{G2} contained 10 people (7 male, 3 female) with severe visual impairments (mean age: $61\pm17$).
Of the latter group, 3 are congenitally blind, while the rest were classified as severely sight impaired later in life.
Of these, 3 participants still have limited light perception with no ability to reliably discern shapes and objects. 
Nevertheless, they were asked to close their eyes during the experiment to. 
None of the participants reported any significant prior experience with electronic navigation aids and none had any hearing or other disabilities that could have influenced their performance in the experiments.
These demographics are summarised in \cref{tab:demographics}.}
Each participant performed three sets of experiments each, with the two characterisation experiments in \cref{sec:participant_characterisation} preceding the final target-search experiment in \cref{sec:target_search}. 
Both groups were given some time before the target-search experiment to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
Furthermore, to minimise any potential speed/accuracy biases, we asked the participants to focus on finding the targets without worrying about the time it took to complete the task.

>>>>

\begin{abstract}
  Sound perception is a fundamental skill for many people with severe sight impairments.
  The research presented in this paper is part of an ongoing project with the aim to create a mobile guidance aid to help people with vision impairments find objects within an unknown indoor environment.
  This system requires an effective non-visual interface and uses bone-conduction headphones to transmit audio instructions to the user.
  It has been implemented and tested with spatialised audio cues, which convey the direction of a predefined target in 3D space.
  We present an in-depth evaluation of the audio interface with several experiments that involve a large number of participants, both blindfolded and with actual visual impairments, and analyse the pros and cons of our design choices.
  In addition to producing results comparable to the state-of-the-art, we found that Fitts's Law (a predictive model for human movement) provides a suitable a metric that can be used to improve and refine the quality of the audio interface in future mobile navigation aids. 
\end{abstract}

@@

\hl{The ActiVis project is an effort to create a fully mobile phone-based navigation aid for people with visual impairments to guide them towards a desired object or location within an unknown indoor environment.
Improvements to computer vision and machine learning techniques, as well as mobile computing hardware performance, are exploited to make this system possible.
In particular, techniques from the active vision field are used to enable a mobile device to gather information on the surrounding environment and use it to generate guidance instructions for a user with limited or no vision.
However, in the literature, these techniques are typically limited to directing electro-mechanical servos~\citep{bajcsy2018revisiting} and in this project, we attempt to find out if similar techniques can be used to direct a user's attention to a target location. 
The proposed system was introduced by~\citet{lock2017portable}, who describe the aforementioned autonomous guidance system paired with a co-adaptive human-machine interface that changes its own parameters over time to better match the user's perception strengths and limitations.
In previous work we developed a prototype guidance system that uses active vision and machine learning models to gather information and help a person find objects within an unknown indoor environment, showing that these techniques can indeed successfully be applied to direct humans' attention~\citep{lock2019active}.
However, in this work we examine the effectiveness of the proposed interface for a searching task and investigate a metric that can be used in the next phases of the project to enable the aforementioned co-adaptive paradigm that could benefit the user experience and boost long-term navigation performance.}

The hardware used for the guidance system prototype are a Google Project Tango device\footnote{https://en.wikipedia.org/wiki/Tango\_(platform)}, along with a set of bone-conduction headphones (both are shown in \cref{fig:participant}).
This Tango device uses a set of embedded RGB-D cameras and other hardware and software components to provide accurate real-time localisation estimates and has powerful image-processing capabilities.
It also gives access to Android's full range of interface and I/O options.
Unlike normal headphones, bone-conduction headphones are placed on and conduct audio signals through a user's cheekbones, instead of their ear canals.
These headphones have the benefit of not interfering with a user's normal hearing function.

@@

The main contributions of this paper are two-fold: 
\begin{itemize}
  \item we provide comprehensive experimental results, with two groups of participants with healthy and limited eyesight, on how well a tone with varying pitch can convey a target's elevation angle when using a mobile device with a bone-conduction headset; 
  \item we show that this sound-based human-machine interface exhibits a response that is well-modelled by Fitts's Law which can provide a useful metric of performance for similar mobile user interfaces.
\end{itemize}

@@

\hl{Over the years, commercial and academic groups have devised new and innovative mobile navigation and electronic travel aids (ETA) for people with visual impairments to address macro- and micro-navigation issues.
The latter refers to the scale in which navigation takes place, with macro-navigation addressing directional guidance on a topographical map using turn-by-turn instructions, for example, while micro-navigation focusses on conveying information of the user's immediate surroundings~\citep{petrie1997mobic}. 
Many of the proposed systems make use of one or a combination of vocal~\citep{mocanu2016when,chessa2016integrated,kanwal2015navigation,sato2019navcog}, audio~\citep{schwarze2015intuitive,rodriguez2012obstacle,kammoun2012navigation} and haptic~\citep{rivera-rubio2015assistive,lee2015rgb,xiao2015assistive} feedback media to present a user with macro- and/or micro-guidance instructions, each of which has its own set of features and limitations.
The interface for this work is used to provide instructions to a stationary user to guide them to point a camera at a target location and we therefore are mainly concerned with micro-navigation tools. 
In this context, respondents with visual impairments typically prefer haptic and vocal feedback over an audio tone~\citep{arditi2013user,golledge2004stated}.
However, from an ergonomic perspective, haptic feedback has significant shortcomings and require additional external hardware to transmit guidance instructions with sufficient resolution.
Furthermore, in a micro-navigation task, high-resolution guidance and many adjustments are required to reach the target location.
In this regard, both haptic and vocal feedback can present a cognitive burden and overwhelm a user's input bandwidth, which could have a detrimental effect on performance and the overall user experience~\citep{klatzky2006cognitive}.}
%Indeed, \citet{arditi2013user} reports that users prefer controlling the rate at which guidance signals are presented to them, instead of being given constant guidance. 
As an alternative, simple audio tones are less prone to these bandwidth and hardware limitations.
However, such tones can potentially fatigue a user if they are too unpleasant.
Researchers have started using cameras and object detectors to determine what a user is looking at and then use simple audio tones and vocal feedback signals to indicate to the user where to find an object.
They report favourable results with these systems~\citep{schauerte2012assistive,tian2013computer,fiannaca2014headlock,vazquez2012helping}.

Work has been done in an attempt to spatialise these tones with a head-related transfer function (HRTF) that simulates a sound source being placed at some arbitrary 3D position.
The authors generally report favourable results when using normal over-ear headphones or speakers~\citep{geronazzo2016interactive,wilson2007swan,katz2010navig,blum2013spatialized,presti2019watchout,crispien1994guib}.
However, the audio transmission device used has a significant effect on performance.
Indeed, research has shown that cheaper headphones and bone-conduction headphones have diminished performance in location perception with HRTFs when compared to over-ear or other expensive headphones~\citep{stanley2006lateralization,voong2019auditory,mascetti2016sonification}.
However, this performance degradation seems limited to the elevation dimension, which is drastically improved when the HRTFs are adjusted for the bone-conduction pathway~\citep{stanley2006lateralization,pec2008individual}.
Another way around the performance issue in the elevation dimension is to transmit the elevation angle by adjusting the audio tone's pitch, such as in the work by~\citet{durette2008visuo}.
We expanded upon the latter's work and investigated using spatialised audio and a varying pitch for guidance in an initial study~\cite{lock2019bone}, and found that participants are able to adequately determine a target's elevation.
The measured performance is comparable to those of more expensive and over-ear headphones.

Fitts's Law~\citep{fitts1954information} is a predictive model of human movement and is particularly useful to evaluate human-computer interactions.
Indeed, researchers have previously used Fitts's Law, and more recently MacKenzie's modified version of it~\citep{mackenzie1992fitts}, as a metric to evaluate the performance of a spatial audio HMI\@.
Fitts's Law was originally proposed for visual target search tasks, but has since been applied in non-visual target search tasks as well.
For example, experiments with a haptic feedback pointing device have been performed to evaluate how effective it was at directing a user towards a target~\citep{ahmaniemi2009augmented} and the authors showed that the search time adheres to Fitts's Law.
However, they also note that it is not a perfect fit, citing the fact that Fitts's Law does not take into account a user's search strategy.
Another group of researchers conducted experiments using a spatial audio interface to describe the position of a target on the horizontal plane~\citep{marentakis2006effects}.
Here, participants pointed to where they thought the targets were, on their left or right, as they traversed a path.
Their results show a good relation between target difficulty and search time, providing a strong argument that Fitts's Law can be used to describe the performance of a spatial audio interface.
These results have since been supported by other authors, who found that Fitts's Law provided a good explanation for the results from an experiment using visual, limited visual and non-visual feedback cues~\citep{wu2010fitts}.
\hl{However, it is not clear whether Fitts's Law applies to a spatial tone that uses varying pitch to convey the target's elevation angle, as demonstrated in this paper.}

@@

\hl{Existing electronic navigation aids have typically struggled to gain significant market traction and replace the traditional walking cane as the standard assistive tool for people with visual impairments.
Current technological limitations include prohibitive costs, bulky hardware requirements and non-user-friendly interfaces~\citep{golledge2004stated,yusif2016older,arditi2013user}.
To address these issues, we implemented a handheld mobile system that is based on a concept proposed by~\citet{lock2019active} and tested by~\citet{lock2019bone} using a Google Tango device that is able to localise itself in real-time.
This system has the benefit of minimal hardware requirements and a compact, familiar form-factor, which will help to overcome the hurdle of user-acceptance and usability.}

\hl{People with visual impairments rely heavily on their hearing~\citep{golledge2004stated} and we wish to avoid blocking this information pathway, since doing so can have severe and undesirable effects~\citep{lichtenstein2012headphone}.
We therefore use a set of bone-conduction headphones that are placed on a user's cheekbones and conduct audio signals through the skull to the inner ear, completely avoiding the outer ear and does therefore does not impede the user's ambient sound perception abilities. 
Alternative, open-back headphones were also considered, but it was found that they still interfere with hearing and were therefore disregarded.
The AfterShockz headphones (shown in \cref{fig:participant}) were ultimately selected.
These headphones have the added benefit of a more discreet form-factor when compared to other over-headphones, thereby addressing the issue of user acceptance.}

@@

\hl{Humans are able to localise sounds in three dimensions by extracting different cues from audio signals~\citep{blauert1969sound,blauert1997spatial}.
These include binaural, where the user compares the signals received at both ears (e.g.\ inter-aural time and level differences), and monaural cues, where a cue is extracted from the signal received at each ear (e.g.\ the spectral profile, audio intensity).
%Binaural cues allow a user to localise a sound on the horizontal plane (e.g.\ a sound from the right will hit the right ear before left one), while monaural cues indicate a sound's distance (e.g. louder sounds appear closer than softer ones) and its position on the vertical plane (e.g.\ the spectral profile generated from the signal hitting the user's body).
Prior to transmitting an audio signal to the user, it can be adjusted by an HRTF to mimic a natural sound source and make a user believe it is located at some arbitrary 3D location. 
An HRTF is a mathematical function that simulates a human head and ear's response to an external sound and is derived by capturing key characteristics that affect the monaural and binaural responses at the user's ear, such as the hearing level and anatomy. 
An HRTF will produce the most accurate results when it is customised to match a specific user, since each user's hearing response is unique, but this is a complex process and HRTFs generated with average values (head sizes, heights, ear shapes, etc.) produce acceptable results~\citep{gardner1995hrtf}.}

\hl{The interface presents the user with guidance instructions in the form of angular adjustments that are required in the pan and elevation dimensions to point a camera at the desired target (see \cref{fig:participant}).
Spatialised audio is well-suited to this micro-navigation task, displaying accuracy comparable to vocal feedback, but with less cognitive load when used in a high-resolution search task such as this~\citep{klatzky2006cognitive}. 
However, we propose a linear adjustment to the signal's pitch as a function of the elevation angle to overcome the limitations of bone-conduction headphones and spatialised audio stated previously.
The pan angle can still be conveyed with a spatialised audio signal generated by an HRTF.
Indeed, this dimension is unaffected by the use of bone-conduction headphones~\citep{schonstein2008comparison,macdonald2006spatial,stanley2006lateralization,lock2019bone}. 
This interface was implemented and evaluated in~\citet{lock2019bone}.}

@@

\hl{To localise a sound on the horizontal plane, the human audition system compares characteristics from the signals received at both ears (binaural cues), such as their volume difference and the time delay between the same sound reaching both ears~\citep{blauert1969sound}.
For example, a sound placed at a person's right will hit their right ear first with a slightly higher volume compared to the left ear.
Since the binaural cues are largely independent if the signal's spectral profile, it would be convenient to use a simple audio wave to transmit guidance instructions to the user.
Therefore, in this work a pure sine wave was used, but can easily be replaced with a richer tone once the system's characteristics are better understood. 
The since was was spatialised using OpenAL's default HRTF, based on the MIT's KEMAR dataset~\citep{hiebert2005openal}.} %which uses the person and targets' positions as input, and outputs a transformed audio signal.}

@@

\hl{To compensate for the loss of elevation localisation performance when conveying spatialised audio via bone-conduction headphones~\citep{macdonald2006spatial,schonstein2008comparison}, we adjust the sine wave's frequency (i.e.\ audible pitch) as a log-linear function of the target's elevation angle, as shown in \cref{fig:pitch-gain}.
When the required elevation adjustment is above or below where the camera is currently pointed, the pitch is increased or lowered respectively. 
This high/low scheme was selected based on humans' natural association of high-pitched sounds with elevated objects and vice-versa for low-lying objects~\citep{pratt1930spatial,blauert1997spatial}.
The pitch is constantly adjusted and updated at \SI{10}{\hertz} in octave- and semitone-based intervals to ensure perceptible changes, while keeping the tone's timbre constant~\citep{shepard1964circularity}. }

\hl{The pitch is changed as a linear function of the elevation angle, the gradient of which is determined by setting the angle and pitch limits.
These limits are set at some number of octaves from the neutral pitch that is emitted when the camera is on-target. 
This tone is heuristically set to \SI{512}{\hertz} following practical tests, which is comfortably audible for a large number of octaves changes.
For this work, we only consider a \SI{180}{\degree} field of view in front of the user and limit the elevation angle to a range of \SI{\pm90}{\degree}, or $[-\frac{\pi}{2}, \frac{\pi}{2}]$.
In previous tests with this specific interface, its performance was comparable to normal over-ear headphones transmitting a signal fully spatialised in the pan and elevation dimensions~\citep{lock2019bone}.}

@@

We performed a set of experiments with the audio interface to determine how effective it is at directing a user to adjust the pan and elevation angles of a camera to point it to a target.
Furthermore, we also carried out a set of pre-screening experiments to determine each participant's hearing characteristics in order to determine their perception limits in the respective audio dimensions.
The participants were given time before the experiments commenced to familiarise themselves with the device and the tones it emits, as well as what the `on-target' tone sounds like.
We also tested the system with a group of participants with severe sight impairments and compared their data to the blindfolded participant dataset.
The results from the experiments we performed allow us to better understand how the users respond to different settings for the spatial audio feedback stimulus and use them to improve and optimise the behaviour of the feedback modes in our portable navigation aid.

@@

\hl{Even though the Tango is able to detect distances to object, we opted to use virtual targets given the added ability to place targets at any random location without physical manipulation without losing the ability to determine its exact position relative to the device.
When this interface is used with a real navigation system, it will be extended to work with real targets and objects.}

@@

A set of experiments were conducted to determine the interface's guidance effectiveness for a pointing task.
These experiments captured the difference between the targets' actual directions and the directions the participants perceived them to be located. 
The participants were given a Tango tablet running an app implementing the experimental setup in \cref{fig:pipeline}.
This app generates a set of virtual targets, one at a time, and presents their directions to the participants with audio guidance signals.
These targets are generated at a constant distance from the participants with pan and elevation angles that are uniformly distributed across the four quadrants to avoid clustering.
Each target's relative angular position is communicated to the participants in real-time as the device is moved around via the bone-conduction headphones.
Every time a participant was confident that they were pointing at a target, i.e.\ hearing a signal that is placed at their front at a frequency of \SI{512}{\hertz}, they tapped the device's screen, marking the location and generating a new target.
The targets are all positioned relative to the device's camera coordinate system, which is tracked by the Tango hardware and localisation API.\
Each participant was tasked with finding 28 targets each.

@@

We also compare the performance of the three pitch gradient settings in terms of the time it takes each participant to find a target.
However, since each participant was presented with a different, randomly generated set of targets, a direct time comparison is not possible.
\hl{Instead, we use Fitts's Law~\citep{fitts1954information}, modified by MacKenzie for uncertain target sizes and noisy data~\citep{mackenzie1992fitts}, which states that there is a relation between the time it takes to find a target and its index of difficulty ($ID$, the ratio between the distance to the target and its width).
It also provides a so-called `index of performance' ($IP$), which can be used to compare the results between the three different configurations used in the experiments.
Furthermore, if a Fitts-like relationship is indeed found, it can be used as an optimisation metric in the co-adaptive interface described by~\citet{lock2017portable}, where interface parameters (e.g.\ pitch gradient, volume setting, etc.) are automatically tweaked in order to maximise target finding performance.}

@@

\noindent \hl{as proposed by~\citet{wobbrock2011effects}, taken as the angle between the participant's target selection and target's actual angular position.
The virtual targets have a programmed radius of approximately} \SI{10}{\centi\metre} (equivalent to approximately \SI{0.1}{\radian}) and when the participant pointed the camera within this radius, \SI{512}{\hertz} on-target tone was emitted.
\hl{This is similar to the work by}~\citet{kabbash1995prince}\hl{, where they had participants search for a point-target with an oversized cursor and elicited a Fitts-like response.
However, in this case the participants were not explicitly informed when they were within this radius and had to use their own subjective judgement of the audio tone.
Since the judgement accuracy will vary across different participants, effectively giving different target sizes, it was necessary to use the effective width as an approximation for the actual target size.
Previous authors have found this to capture the true performance metrics fairly well}~\citep{zhai2004speed}.
Fitts's index of performance, $IP$, can then be calculated using

@@

Two groups of participants were recruited for the experiments on a volunteer basis. 
\hl{Group~\textit{G1} consisted of 42 undergraduate students (10 male, 32, female) with normal eyesight who were blindfolded for the experiments (mean age: $20\pm2$ years).
Group \textit{G2} contained 10 people (7 male, 3 female) with severe visual impairments (mean age: $61\pm17$ years).
Of the latter group, 3 are congenitally blind, while the rest were classified as severely sight impaired later in life.
Of these, 3 participants still have limited light perception with no ability to reliably discern shapes and objects (the rest had no light perception). 
Nevertheless, they were asked to close their eyes during the experiment. 
None of the participants reported any significant prior experience with electronic navigation aids and none had any hearing or other disabilities that could have influenced their performance in the experiments.
These demographics are summarised in \cref{tab:demographics}.}

Each participant performed three sets of experiments each, with the two characterisation experiments in \cref{sec:participant_characterisation} preceding the final target-search experiment in \cref{sec:target_search}. 
Both groups were given some time before the target-search experiment to familiarise themselves with the system, the audio signal's behaviour and the \SI{512}{\hertz} on-level tone. 
Furthermore, to minimise any potential speed/accuracy biases, we asked the participants to focus on finding the targets without worrying about the time it took to complete the task.
